---
title: "p8105_hw1_rj2543"
author: "rj2543"
date: "2018/9/15"
output: github_document
---

# Problem 1

First, create a data frame comprised of:

* A random sample of size 10 from a uniform[0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector

```{r 1create}
library(tidyverse)
problem1_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = vec_numeric > 2,
  vec_char = character(length = 10),
  vec_factor = factor(x = vec_char)
)
```

Try to take the mean of each variable in the dataframe. 

```{r 1mean}
mean(problem1_df$vec_numeric)
mean(problem1_df$vec_logical)
mean(problem1_df$vec_char)
mean(problem1_df$vec_factor)
```

The **mean** function only works for the *sample* and the *logical vector*, because in the explanation of **mean** function, it is said that if x is not logical (coerced to numeric), numeric (including integer) or complex, NA_real_ is returned, with a warning.

In some cases, one can explicitly convert variables from one type to another.

```{r 1convert1code, eval = FALSE}
as.numeric(problem1_df$vec_logical)
as.numeric(problem1_df$vec_char)
as.numeric(problem1_df$vec_factor)
```

In this case, what happens?

```{r 1convert1result, echo = FALSE}
as.numeric(problem1_df$vec_logical)
as.numeric(problem1_df$vec_char)
as.numeric(problem1_df$vec_factor)
```


Next, try to convert the character variable from character to factor to numeric; similarly, convert the factor variable from factor to character to numeric. 
What happens?

```{r 1convert2}
as.numeric(as.factor(problem1_df$vec_char))
as.numeric(as.character(problem1_df$vec_factor))
```


