---
title: "p8105_hw1_rj2543"
author: "rj2543"
date: "2018/9/15"
output: github_document
---

# Problem 1

This problem focuses on variable types and coercion.

First, create a data frame comprised of:

* A random sample of size 10 from a uniform[0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector

```{r 1create}
library(tidyverse)
problem1_df = tibble(
  sample_numeric = runif(10, min = 0, max = 5),
  sample_logical = (sample_numeric > 2),
  sample_char = character(length = 10),
  sample_factor = factor(x = sample_char)
)
```

Try to take the mean of each variable in the dataframe. 

```{r 1mean}
mean(problem1_df$sample_numeric)
mean(problem1_df$sample_logical)
mean(problem1_df$sample_char)
mean(problem1_df$sample_factor)
```

The **mean** function only works for the *numeric* and the *logical*, because in the explanation of **mean** function, it is said that if x is not logical (coerced to numeric), numeric (including integer) or complex, NA_real_ is returned, with a warning.

In some cases, one can explicitly convert variables from one type to another.

```{r 1convert1code, eval = FALSE}
as.numeric(problem1_df$sample_logical)
as.numeric(problem1_df$sample_char)
as.numeric(problem1_df$sample_factor)
```

In this case, what happens?

```{r 1convert1result, echo = FALSE}
as.numeric(problem1_df$sample_logical)
as.numeric(problem1_df$sample_char)
as.numeric(problem1_df$sample_factor)
```


Next, try to convert the character variable from character to factor to numeric; similarly, convert the factor variable from factor to character to numeric. 
What happens?

```{r 1convert2}
as.numeric(as.factor(problem1_df$sample_char))
as.numeric(as.character(problem1_df$sample_factor))
```

# Problem 2

This problem focuses on plotting and the use of inline R code.

Create a data frame comprised of:

* x: a random sample of size 1000 from a standard Normal distribution
* y: a random sample of size 1000 from a standard Normal distribution
* A logical vector indicating whether the x + y > 0
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector

```{r 2create}
problem2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  xy_logical = ((x+y) > 0),
  xy_lognumeric = as.numeric(xy_logical),
  xy_logfactor = as.factor(xy_logical)
)
```

The short description includes:

* the size of the dataset is **`r length(problem2_df)`**
* the mean of x is **`r mean(problem2_df$x)`** and the median of x is **`r median(problem2_df$x)`**
* the proportion of cases for which the logical vector is TRUE is **`r sum(problem2_df$xy_lognumeric)/length(problem2_df$xy_logical)`**

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

```{r 2plot}
ggplot(problem2_df, aes(x = x, y = y, color = xy_logical)) + geom_point()
ggplot(problem2_df, aes(x = x, y = y, color = xy_lognumeric)) + geom_point()
ggplot(problem2_df, aes(x = x, y = y, color = xy_logfactor)) + geom_point()
```

Export the first scatterplot to the project directory using ggsave.

```{r 2save}
ggsave('scatterplot of y vs x colored by logical.pdf', plot = ggplot(problem2_df, aes(x = x, y = y, color = xy_logical)) + geom_point())
```

